---
title: "basilisk and owlready2 -- some demonstrations"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{littlebas}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::pdf_document:
    toc: yes
    number_sections: yes
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction 

We will look at the classes returned by
an owlready2 load event.  The littlebas
package includes `get_classes`.  What
can we do with a python iterator in R
to learn about an ontology?

# The basic source code

We are getting owlready2 from pip with a specific
version.  We also dictate that we will use python 3.9

```
# necessary for python module control
bsklenv <- basilisk::BasiliskEnvironment(envname="bsklenv",
    pkgname="littlebas", packages="python=3.9", pip="owlready2==0.45")

#' demo
#' @param owlfile path to owl file
#' @export
get_classes = function(owlfile) {
 proc = basilisk::basiliskStart(bsklenv)
 basilisk::basiliskRun(proc, function(owlfile) {
 o2 = reticulate::import("owlready2") # 'cached' by reticulate?
 o2$get_ontology(owlfile)$load()$classes()
 }, owlfile)
}
```

# Run it

We include a copy of cl.owl, compressed, in the
littlebas package.  With ontoProc such a file would be in
BiocFileCache.

```{r doit}
library(littlebas)
cloppath = R.utils::decompressFile(
   system.file("owl", "cl.owl.gz", package="littlebas"), 
   temporary=TRUE, ext="gz", FUN=gzfile, remove=FALSE,
   overwrite=TRUE)
clc = get_classes(cloppath)
clc
```

Now we have an instance of a python generator.  Use `py_help`
from reticulate to learn about it.

```
Help on generator object:

classes = class generator(object)
 |  Methods defined here:
 |  
 |  __del__(...)
 |  
 |  __getattribute__(self, name, /)
 |      Return getattr(self, name).
 |  
 |  __iter__(self, /)
 |      Implement iter(self).
 |  
 |  __next__(self, /)
 |      Implement next(self).
...
```

We can just try to use the "next" method.
```{r tryit}
clc$`__next__`()
clc$`__next__`()
c3 = clc$`__next__`()
class(c3)
names(c3)
```

So there is a lot of information in each output of
the generator.  Let's work with 1000 steps.
We'll start over.

```{r doagain}
clc$close()
clc = get_classes(cloppath)
stuff = lapply(1:1000, function(x) clc$`__next__`())
stuff[[3]]
stuff[[1000]]
stuff[[1000]]$label
stuff[[1000]]$label[0]
stuff[[1000]]$name
```

Seems super fast to iterate through the ontology generator.
We can collect labels and terms using code like that above.

How do we know how far to go?  Try a while loop with
a try call to check for end of generation.
